---
title: "Statistical inference On Education attained based on Social conditions"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(reshape2)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `gss`. Delete this note when before you submit 
your work. 

```{r load-data}
load("gss.Rdata")
```



* * *

## Part 1: Data
```{r Data}
#str(gss)
colnames(gss)[colSums(is.na(gss))==0]
colSums(is.na(gss))
#Filter data based on reauired fields according to the Reasearch question.
df2<-gss[c("sex","race","educ")]
str(df2)
#NA values are a small proportion, hence delete the corresponding rows
df2<-df2[complete.cases(df2),]
str(df2)
```

* * *

## Part 2: Research question
Research Question: Is "Highest year of attending school" dependent on Gender and Race ? <br />
This is a quesion of Social importance as this would determine any Discrimination/conditions effecting people's Schooling based on Social Status like Gender and Race. 


* * *

## Part 3: Exploratory data analysis
```{r Exploratory Data Analysis}

summary(df2)

ggplot(df2,aes(x=df2$educ))+geom_histogram()+labs(x="Highest year of school completed",y="Frequency",title="Frequency distribution of Highest year of school attended")

boxplot(df2$educ~df2$sex,main="Distribution of Highest level Education across Gender",xlab="Gender",ylab = "Highest Education")
boxplot(df2$educ~df2$race,main="Distribution of Highest level Education across Race",xlab = "Race", ylab = "Highest Education")
col_both<-cbind(df2$race,df2$sex)
ggplot(aes(y = df2$educ, x = df2$race, fill = df2$sex), data = df2) + geom_boxplot()+labs(x="Race and Gender",y="Highest Education",title = "Distribution of Highest level Education across Race and Gender",fill="Gender")


```

```{r by-means}
df2 %>%
  group_by(race,sex) %>%
  summarise(mean_highest_education = mean(educ))
```


* * *

## Part 4: Inference

Null Hypothesis: There is no relationship between Highest education and Gender and race.<br/>
Alternative Hypothesis: There is relationship between Highest education and Gender and race.

```{r inference for Research question}
#Hypothesis test for just Gender vs Highest education achieved
inference(y = educ, x = sex, data = df2, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")

#Chi squared test
count_matrix<-df2 %>%
   group_by(race,sex) %>%
   summarise(sum_highest_education = sum(educ))
count_matrix
#Get a 2d count matrix
cont_matrix<-matrix(data=count_matrix$sum_highest_education,ncol = 2,byrow = TRUE)
colnames(cont_matrix)<-c("male","female")
rownames(cont_matrix)<-c("white","black","other")
cont_matrix
chisq <- chisq.test(cont_matrix)
chisq
chisq$p.value




```
Conclusion: Since p-value is less than significance level(0.05) We reject the Null Hypothesis and conclude that There is a significant relationship between Gender, Race and he Highest education received. 